DATA 605 Final

### 2019-05-16 DATA605/MSilva_Final.Rmd
More Work on DATA605 Final

@@

```{r}
set.seed(12345)
N <- 10
X <- runif(10000, min=1, max=N)
mu <- sigma <- (N + 1)/2
Y <- rnorm(10000, mean=mu, sd=sigma)
```

@@

a. P(X>x | X>y)

b. P(X>x, Y>y)

c. P(X<x | X>y)

@@

Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

```{r, comment=NA}
set.seed(12345)
N <- 10
n <- 10000
mu <- sigma <- (N + 1)/2
df <- data.frame(X = runif(n, min=1, max=N), 
                 Y = rnorm(n, mean=mu, sd=sigma))
summary(df$X)
summary(df$Y)
```

@@

```{r}
x <- median(df$X)
y <- as.numeric(quantile(df$Y)["25%"])
```

small letter "x" is `r x` and small letter "y" is `r y`.

a. $P(X>x | X>y)$

```{r}
P_a_and_b <- df %>%
  filter(X > x,
         X > y) %>%
  nrow() / n

P_b <- df %>%
  filter(X > y) %>%
  nrow() / n

problem_1a <- P_a_and_b / P_b
```

**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` given that it is greater than `r y` is `r problem_1a`**

b. $P(X>x, Y>y)$

```{r}
problem_1b <- df %>%
  filter(X > x,
         Y > y) %>%
  nrow() / n
```

**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` and a random normally distributed number with a mean and standard deviation of `r mu` being greater than `r y` is `r problem_1b`**

c. $P(X<x | X>y)$

```{r}
P_a_and_b <- df %>%
  filter(X < x,
         X > y) %>%
  nrow() / n

P_b <- df %>%
  filter(X > y) %>%
  nrow() / n

problem_1c <- P_a_and_b / P_b
```

**The probablity of a random number uniformly ranging from 1 to `r N` being less than `r x` given that it is greater than `r y` is `r problem_1c`**

@@

```{r}
# Create Joint Probabilities
temp <- df %>%
  mutate(A = ifelse(X > x, " X greater than x", " X not greater than x"),
         B = ifelse(Y > y, " Y greater than y", " Y not greater than y")) %>%
  group_by(A, B) %>%
  summarise(count = n()) %>%
  mutate(probability = count / n)

# Create Marginal Probabilities
temp <- temp %>%
  ungroup() %>%
  group_by(A) %>%
  summarise(count = sum(count),
            probability = sum(probability)) %>%
  mutate(B = "Total") %>%
  bind_rows(temp)

temp <- temp %>%
  ungroup() %>%
  group_by(B) %>%
  summarise(count = sum(count),
            probability = sum(probability)) %>%
  mutate(A = "Total") %>%
  bind_rows(temp)

# Create Table
temp %>%
  select(-count) %>%
  spread(A, probability) %>%
  rename(" " = B) %>%
  kable() %>%
  kable_styling()
```

**P(X>x and Y>y) is `r temp[temp$A == " X greater than x" & temp$B == " Y greater than y",]$probability`. P(X>x)P(Y>y) is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability` $\times$ `r temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability` which is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability * temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability`.  They are not the same.**

Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?

```{r, comment=NA}
count_data <- temp %>%
  filter(A != "Total",
         B != "Total") %>%
  select(-probability) %>%
  spread(A, count) %>%
  as.data.frame()

row.names(count_data) <- count_data$B

count_data <- count_data %>%
  select(-B) %>%
  as.matrix()

fisher.test(count_data)
```

```{r, comment=NA}
chisq.test(count_data)
```

**Fisher's Exact Test is for is used when you have small cell sizes (less than 5).  The Chi Square Test is used when the cell sizes are large.  It would be appropriate in this case.**
